---
title: "Assignment -1 24440622"
output:
  pdf_document: default
  html_document: default
date: "2026-07-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Objectives:

- Online retailers has client with heavy buying and different structural patterns.They are having important strategic values.By understanding which items frequently co-purchased client can bundle [...]
- Items that exhibit strong positive cross-correlations should be dynamically recommended on product pages ("Frequently bought together") to cross-sell effortlessly. Wholesalers purchase predictab[...]
- Identifying rules allows the retailer to send highly personalized, volume-tiered email promotions tailored to the specific operational niches of those wholesalers

```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(arules)
library(arulesViz)
```

## 2. Data set description

- The raw data set contains $541,909$ rows covering international transactions.

```{r,warning=FALSE}
#Loading raw data set
raw_data <- read_csv("Online_Retail.csv", col_types = cols(
  InvoiceNo = col_character(),
  StockCode = col_character(),
  Description = col_character(),
  Quantity = col_integer(),
  InvoiceDate = col_datetime(format = "%d/%m/%Y %H:%M"),
  UnitPrice = col_double(),
  CustomerID = col_character(),
  Country = col_character()))
```

```{r}
head(raw_data)
```

### Pre processing

- Filter by geographic region - United Kingdom
- Removed cancelled transactions
- Removed non product codes
- Handling missing values

```{r}
#Preprocessing
cleaned_data <- raw_data %>% 
  filter(Country == "United Kingdom") %>% 
  filter(!is.na(CustomerID) & !is.na(Description)) %>%
  filter(!str_detect(InvoiceNo, "^[Cc]") & Quantity > 0 & UnitPrice > 0) %>%
  filter(!StockCode %in% c("POST", "DOT", "M", "BANK CHARGES", "PADS")) %>%
  mutate(Description = str_trim(Description))
```

```{r}
head(cleaned_data)
```

- Aggregated to basket format

```{r}
basket_data <- cleaned_data %>% group_by(InvoiceNo) %>% summarise(Items = paste(Description, collapse = ","))
head(basket_data)
```

```{r}
write_lines(basket_data$Items, "market_basket_transactions.csv")
```

### Rule Mining Process

Exploration Cycles and Parameter Settings

- Cycle 1 (Exploratory): Initial configurations are set broadly with support = 0.005 (items appearing in at least 0.5% of baskets) and confidence = 0.5 (50% chance of co-purchase). This serves to [...]

- Cycle 2 (Refinement): To sift out the noise, thresholds are tuned to support = 0.01 and confidence = 0.6, prioritizing stronger relationships that carry higher operational certainty.

```{r,warning=FALSE}
transactions <- read.transactions("market_basket_transactions.csv", 
                                  format = "basket", 
                                  sep = ",",
                                  rm.duplicates = TRUE)
```

```{r}
print(summary(transactions))
```

```{r}
itemFrequencyPlot(transactions, topN = 20, type = "relative", 
                  main = "Top 20 Most Frequently Purchased Items (UK)")
```

```{r}
rules <- apriori(transactions, 
                 parameter = list(supp = 0.01, conf = 0.6, minlen = 2))
```

```{r}
cat(paste("\nTotal association rules discovered:", length(rules), "\n"))
```

- Redundancy Pruning: Rules that are subsets of broader, more impactful rules with a lower lift are automatically pruned out to ensure that the resulting rules contain distinct, actionable insigh[...]

```{r}
rules_sorted <- sort(rules, by = "lift", decreasing = TRUE)
```

```{r}
subset_matrix <- is.subset(rules_sorted, rules_sorted, proper = TRUE)
subset_matrix[lower.tri(subset_matrix, diag = TRUE)] <- FALSE
redundant <- colSums(subset_matrix, na.rm = TRUE) >= 1
rules_pruned <- rules_sorted[!redundant]
```

```{r}
inspect(rules_pruned[1:10])
```

## Resulting Rules

After pruning, the algorithm surfaces tight, high-lift association rules. The metrics evaluated are:

- Support: The absolute popularity of the rule.
- Confidence: The predictive reliability of the rule.
- Lift: The strength of the rule over random coincidence (values $>1$ indicate strong positive dependency).

```{r}
write(rules_pruned, file = "discovered_association_rules.csv", 
      sep = ",", row.names = FALSE)
```

## Recommendations

Based on the high-yield rules uncovered, the client should implement the following

- strategic changes:Strategic Virtual Bundling (Cross-Category Pairs): Create a "Regency High-Tea Set" or "Gardener's Comfort Set" containing the highly correlated items bundled together. Offer a[...]

- Dynamic Recommendations and UI Design: The web development team should hard-code dynamic recommendation blocks. When a customer adds the GREEN REGENCY TEACUP to their virtual cart, an interstit[...]

- Inventory Allocation Syncing: Tie warehouse inventory alerts together for these coupled items. Because these products sell in tandem, experiencing an out-of-stock scenario on the ALARM CLOCK BA[...]
